Traffic Aware Resource Allocation Schemes for Multi-Cell \acs{MIMO}-\acs{OFDM} Systems

### 2015-03-11 Report/sub_channel_wsrm.tex
revised
updated nam_comments without appendix

The complexity of the \ac{JSFRA} algorithm scales with the number of sub-channels considered. In addition, the iterations required for the algorithm convergence increases with the problem size. To overcome the complexity, we can decouple the problem to sub-channel specific subproblems by using decomposition approaches presented in \cite{palomar2006tutorial,boyd2011distributed} to design the precoders using primal or dual decomposition.

As an alternative sub-optimal solution, we present \acl{QM} \ac{SRA}, which solves for the precoders using \ac{JSFRA} formulation only for a specific sub-channel \me{i} with a fixed transmit power \me{P_{\max,i}}. The sharing of power can be equal or based on a predetermined pattern as in partial frequency reuse for the sub-channels by adhering to 
\begin{equation}
\sum_{i=1}^N P_{\max,i} = P_{\max}.
\end{equation}
Even though \me{N} sub-channels are present at any given scheduling instant, precoders are solved for each sub-channel in a sequential manner with the sub-channel specific total power constraint \me{P_{\max,i}} and the backlogged packets. Let \me{Q_{k,i}} be the number of backlogged packets associated with user \me{k} before solving for the precoders specific to the sub-channel \me{i}. Since the precoder design is sequential, \textit{i.e}, the precoders are designed for sub-channels \me{[0,i-1]} before \me{\ith{i}} sub-channel, the number of backlogged packets for the initial sub-channel is initialized as \me{Q_{k,1} = Q_k}. The queues associated with the consecutive sub-channels are given by
\begin{equation}	\label{eqn-weight}
	Q_{k,i+1} = \max{\left \lbrace Q_k - \sum_{j = 1}^{i} \, \sum_{l = 1}^{L} \, t_{l,k,j} ,0 \right \rbrace }, \; \forall \; k \in \mathcal{U},
\end{equation}
where \me{t_{l,k,j}} denotes the rate corresponding to the user \me{k} on the \me{\ith{j}} sub-channel and \me{\ith{l}} spatial stream. Note that the proposed scheme is sensitive to the order of the sub-channel selection due to the sequential precoder design for each sub-channel. However, the \ac{SRA} approach provides faster convergence in contrast to the \ac{JSFRA} formulation due to the substantial reduction in the optimization variables for each sub-channel problem. As the number of user increases, the \ac{SRA} formulation will be insusceptible to the sub-channel ordering due to the multi-user diversity.
%The algorithmic representation of the \ac{SRA} scheme is shown in Algorithm \ref{algo-5}
%\begin{algorithm}
% \SetAlgoLined
% \DontPrintSemicolon
% \BlankLine
% \SetKwInput{KwInit}{Initialize}
% \KwIn{\me{a_k, \, Q_k, \, \mvec{H}{b,k,n},\; \fall b \in \mathcal{B}, \, \fall k \in \mathcal{U}}}
% \KwIn{permute \me{\mathcal{N} \rightarrow \tilde{\mathcal{N}}}}
% \For{\me{n \leftarrow 1 } \KwTo \me{N}}{
% update \me{Q_{k,n}} using \eqref{eqn-weight} and let \me{\hat{n} = \tilde{\mathcal{N}}(n)}\;
% \KwOut{\me{\mvec{m}{l,k,\hat{n}}} and \me{\mvec{w}{l,k,\hat{n}} \fall l \in \set{1,2,\dotsc,L}}}
% \KwInit{\me{i=0} and \me{\tilde{\mbf{m}}_{l,k,\hat{n}}} randomly satisfying per sub-channel power constraint \me{P_{\max,\hat{n}}}}
% update \me{\mvec{w}{l,k,\hat{n}}} and \me{\tilde{\mbf{u}}_{l,k,\hat{n}}} using \eqref{eqn-10} and \eqref{eqn-8}\;
% \Repeat{Queue convergence or \me{i \geq I_{\max}}}{
% initialize \me{j = 0}\;
% \Repeat{\ac{SCA} convergence or \me{j \geq J_{\max}}}{
% solve for \me{\mvec{m}{l,k,\hat{n}}} using \eqref{eqn-9} with per sub-channel power constraint \me{P_{\max,\hat{n}}}\;
% update the constraint set \eqref{eqn-8} with \me{\tilde{u}_{l,k,\hat{n}}} and \me{\mvec{m}{l,k,\hat{n}}} using \eqref{eqn-wsrm-expr}\;
% $j = j + 1$\;
% }
% update \me{\mvec{w}{l,k,n}} using \eqref{eqn-10} with the updated precoders \me{\mvec{m}{l,k,n}}\;
% $i = i + 1$\;
% }
% }
% \caption{Algorithm of \acs{SRA} scheme}
% \label{algo-5}
%\end{algorithm}
%

>>>>

The complexity of the \ac{JSFRA} algorithm scales quickly with the number of sub-channels, since the complexity of the interior point method, which is used to solve the problem, increases quickly with the problem size. Thus, we can use the decomposition methods presented in \cite{palomar2006tutorial,boyd2011distributed} to overcome this complexity by designing precoders for each sub-channels independently with minimal information exchange.

As an alternative sub-optimal solution, we present \acl{QM} \ac{SRA}, which solves for the precoders using \ac{JSFRA} formulation only for a specific sub-channel \me{i} with a fixed transmit power \me{P_{\max,i}}. The sharing of power can be equal or based on a predetermined pattern as in partial frequency reuse for the sub-channels as given by
\begin{equation}
\sum_{i=1}^N P_{\max,i} = P_{\max}.
\end{equation}
Even though \me{N} sub-channels are present at any given scheduling instant, precoders are computed for each sub-channel in a sequential manner with the sub-channel specific total power constraint \me{P_{\max,i}} and the backlogged packets. Let \me{Q_{k,i}} be the number of backlogged packets associated with user \me{k} before solving for the precoders specific to the sub-channel \me{i}. Since the precoder design is sequential, \textit{i.e}, the precoders are designed for sub-channels \me{[0,i-1]} before \me{\ith{i}} sub-channel, the number of backlogged packets for the initial sub-channel is initialized as \me{Q_{k,1} = Q_k}. The queues associated with the consecutive sub-channels are given by
\begin{equation}	\label{eqn-weight}
	Q_{k,i+1} = \max \Big ( Q_k - \sum_{j = 1}^{i} \, \sum_{l = 1}^{L} \, t_{l,k,j} ,0 \Big ) \; \forall \; k \in \mathcal{U}
\end{equation}
where \me{t_{l,k,j}} denotes the rate corresponding to the user \me{k} on the \me{\ith{j}} sub-channel and \me{\ith{l}} spatial stream. Note that the proposed scheme is sensitive to the order of the sub-channel selection due to the sequential precoder design for each sub-channel. However, the \ac{SRA} approach provides faster convergence in contrast to the \ac{JSFRA} formulation due to the substantial reduction in the optimization variables for each sub-channel problem. As the user count increases, the \ac{SRA} formulation will be insusceptible to the sub-channel ordering due to the multi-user diversity.
%The algorithmic representation of the \ac{SRA} scheme is shown in Algorithm \ref{algo-5}
%\begin{algorithm}
% \SetAlgoLined
% \DontPrintSemicolon
% \BlankLine
% \SetKwInput{KwInit}{Initialize}
% \KwIn{\me{a_k, \, Q_k, \, \mvec{H}{b,k,n},\; \fall b \in \mathcal{B}, \, \fall k \in \mathcal{U}}}
% \KwIn{permute \me{\mathcal{N} \rightarrow \tilde{\mathcal{N}}}}
% \For{\me{n \leftarrow 1 } \KwTo \me{N}}{
% update \me{Q_{k,n}} using \eqref{eqn-weight} and let \me{\hat{n} = \tilde{\mathcal{N}}(n)}\;
% \KwOut{\me{\mvec{m}{l,k,\hat{n}}} and \me{\mvec{w}{l,k,\hat{n}} \fall l \in \set{1,2,\dotsc,L}}}
% \KwInit{\me{i=0} and \me{\tilde{\mbf{m}}_{l,k,\hat{n}}} randomly satisfying per sub-channel power constraint \me{P_{\max,\hat{n}}}}
% update \me{\mvec{w}{l,k,\hat{n}}} and \me{\tilde{\mbf{u}}_{l,k,\hat{n}}} using \eqref{eqn-10} and \eqref{eqn-8}\;
% \Repeat{Queue convergence or \me{i \geq I_{\max}}}{
% initialize \me{j = 0}\;
% \Repeat{\ac{SCA} convergence or \me{j \geq J_{\max}}}{
% solve for \me{\mvec{m}{l,k,\hat{n}}} using \eqref{eqn-9} with per sub-channel power constraint \me{P_{\max,\hat{n}}}\;
% update the constraint set \eqref{eqn-8} with \me{\tilde{u}_{l,k,\hat{n}}} and \me{\mvec{m}{l,k,\hat{n}}} using \eqref{eqn-wsrm-expr}\;
% $j = j + 1$\;
% }
% update \me{\mvec{w}{l,k,n}} using \eqref{eqn-10} with the updated precoders \me{\mvec{m}{l,k,n}}\;
% $i = i + 1$\;
% }
% }
% \caption{Algorithm of \acs{SRA} scheme}
% \label{algo-5}
%\end{algorithm}
%
